---
output:
  html_document:
    fig_height: 3
    fig_width: 7
---

```{r echo=FALSE}
knitr::read_chunk('blog_code.R')
```

```{r first, cache=FALSE, echo=FALSE, results='hide', message = FALSE,warning=FALSE}
<<main>>
```

```{r cache=FALSE, echo=FALSE,include=FALSE}
options(scipen=1,digits = 1)
```

```{r cache=FALSE, echo=FALSE,include=FALSE}
player1 <- "Ronan O'Gara"
```

<br>
Rugby has faced many controversies and alleged conspiracies. Was rugby [complicit](https://en.wikipedia.org/wiki/1981_South_Africa_rugby_union_tour) in [apartheid](http://www.teara.govt.nz/en/photograph/26666/all-white-all-blacks-1928)? Whatever happened to [Keith Murdoch](http://www.smh.com.au/rugby-union/union-news/the-murdoch-myth-all-blacks-legend-lives-life-of-a-recluse-after-infamous-punch-20111020-1mb2p.html)? Did [Suzie the waitress](https://en.wikipedia.org/wiki/1995_Rugby_World_Cup_Final) really do it?  One ongoing controversy concerns the extent to which international rugby teams "poach" players from other countries. Most of the heat is usually directed toward New Zealand, who are accused of being particularly predatory in their foreign recruitment.^[Here are a few examples: [1](http://www.telegraph.co.uk/sport/rugbyunion/2361788/Ruthless-All-Blacks-harvest-fertile-Island-talent.html), [2](http://www.dailymail.co.uk/sport/rugbyunion/article-2612324/All-Blacks-remember-players-theyve-poached-Pacific-Islands.html), [3](http://www.telegraph.co.uk/sport/rugbyunion/international/newzealand/11113937/New-Zealand-owe-a-debt-to-Samoa-that-can-only-be-repaid-with-a-full-touring-party-for-long-overdue-first-Test-visit.html), [4](http://www.theroar.com.au/2012/10/31/the-problem-with-eligibility-in-international-rugby/), [5](http://www.stuff.co.nz/sport/rugby/opinion/10132272/Gifford-Home-truths-about-Pasifika-poaching), and [6](http://www.stuff.co.nz/13313/All-Blacks-accused-of-Pacific-pillage-again).] This debate among the international rugby public and media has largely relied on casual observations and anecdotes rather than evidence.^[The [New Zealand Herald](http://www.nzherald.co.nz/sport/news/article.cfm?c_id=4&objectid=11278276) provided a notable exception by compiling a dataset of the birthplace of every [All Black](https://en.wikipedia.org/wiki/New_Zealand_national_rugby_union_team) prior to 2014.] In an attempt to bring some sanity to the discussion, I built a dataset of the birthplace and playing statistics of every international rugby player to have played a test match for the 10 "major" nations.^[I scraped the majority of my data from the superb [ESPNscrum](http://www.espnscrum.com) website, which lists statistics for every person to have played a test-match for the major rugby nations. As well as statistics like matches played, points and tries scored, there is also birth information for the majority of players. However, the birthplace information, where available, is limited to the city of birth. I therefore match the birth cities with their respective countries using the [Geopy](https://pypi.python.org/pypi/geopy) geocoding package in Python. This works well for city names that are unique to a particular country, such as [Taumarunui](https://en.wikipedia.org/wiki/Taumarunui). However, in the case of ambiguous city/country pairs, the Geopy package (to the best of my knowledge) returns the city with the highest population, which leads to errors. For example, the Geopy package attributes those players that were born in the town of Hastings, New Zealand as being born in Hastings, England. I therefore also make manual adjustments to the scraped datasets based on my own research. In the case of New Zealand, I use the dataset provided by the [New Zealand Herald](http://www.nzherald.co.nz/sport/news/article.cfm?c_id=4&objectid=11278276) for players making their test debut prior to 2015. <br> I define the ten major nations as those teams in the Six Nations Championship and the Rugby Championship.] You can explore the data and maybe help me fill out a couple missing datapoints [below](#data).

I focus below on players making their debut in the professional era (post 1996) and define a foreign player as someone who was not born in the country for which they played.^[This definition has some shortcomings. For example, [`r player1`](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Ireland.csv#L957), who scored `r d$Points[d$Name==player1]` points for `r d$Team[d$Name==player1]`, is considered a foreign born player because he was born in the United States, despite the  fact that few would question his Irish rugby "credentials".] The following table shows that, on average, `r round(mean(FB$TotalPlayers))` players debuted for each country since 1996, ranging from `r min(FB$TotalPlayers)` for `r FB$Team[which.min(FB$TotalPlayers)]` to `r max(FB$TotalPlayers)` for `r FB$Team[which.max(FB$TotalPlayers)]`.^[I am missing the birth information for `r fin_df2 %>% filter(Team %in% target) %>% select(MissPlayers) %>% sum()` of these players, and they are not included in this table or the rest of the analysis. `r fin_df2$MissPlayers[fin_df2$Team=="Argentina"]` of the missing players are from Argentina, `r fin_df2$MissPlayers[fin_df2$Team=="Ireland"]` from Ireland, `r fin_df2$MissPlayers[fin_df2$Team=="France"]` from France and `r fin_df2$MissPlayers[fin_df2$Team=="Italy"]` from Italy.]

<br>

<h4>
<center>
Player Debuts in the Professional Era (1996 - 2016)
</center>
</h4>


```{r, results='asis',echo=FALSE,warning=FALSE}
library(ascii)
print(ascii(test,include.rownames = TRUE,format='d',
             align = c("l","c","c","c","c","c","c","c","c","c","c")),type='pandoc')
```




Much of the debate assumes an increasing trend in foreign born players. The graph below presents the total number of foreign players making their debut in each year for the 10 major nations. Although there was a surge in the number of foreign players in 2016, there is otherwise no noticeable trend. There were more foreign born players in 2002 than 2015, for example, and the number has generally fluctuated around an average of `r round(mean(tl$Foreign))` foreign born debutants per year, with a low of `r min(tl$Foreign)` in `r tl$Debut[which.min(tl$Foreign)]` and a high of `r max(tl$Foreign)` in `r tl$Debut[which.max(tl$Foreign)]`. 

<br>

<h4>
<center>
Foreign Born Trend
</center>
</h4>


<center>
```{r, results='asis',echo=FALSE,warning=FALSE}
tl_g
```
</center>

<br>

The graph below shows that `r FB$Team[1]` "poaches" the most. Since 1996, `r FB$"Foreign Born (%)"[1] `% of players making their debut were not born in `r FB$Team[1]`. Next is `r FB$Team[2]` with `r FB$"Foreign Born (%)"[2]`%, Ireland with `r FB$"Foreign Born (%)"[3]`%, Australia with `r FB$"Foreign Born (%)"[4]`% and England with `r FB$"Foreign Born (%)"[5]`%. `r FB$Team[length(target)]` poaches the least with just `r FB$"Foreign Born (%)"[length(target)]`% of its players born overseas, followed by `r FB$Team[length(target)-1]` with `r FB$"Foreign Born (%)"[length(target)-1]`%, `r FB$Team[length(target)-2]` with `r FB$"Foreign Born (%)"[length(target)-2]`% and `r FB$Team[length(target)-3]` with `r FB$"Foreign Born (%)"[length(target)-3]`%.

<br>

<h4> 
<center>
Foreign Born Players
</center>
</h4>

<center>
```{r,echo=FALSE,warning=FALSE}
FB_g
```
</center>

<br>

```{r,echo=FALSE,warning=FALSE,results='hide'}
WB <- m1[m1$variable=="Foreign Born Population (%)",] %>% arrange(-value)
```

It shouldn't be surprising that every team has foreign born players. International migration is a reality of the modern world, and many of the top rugby countries have large foreign born populations. The following graph compares the foreign born composition of each team to the general population in that country.^[I use 2015 census data retrieved from the [World Bank Data Bank](http://databank.worldbank.org/data/reports.aspx?source=2&series=SM.POP.TOTL.ZS&country=#) (accessed on 12 March 2017) for foreign born populations. The dataset is [here](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/WorldBankData/Data.csv) and details of the definitions and sources for each country can be found [here](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/WorldBankData/DefinitionandSource.csv). I do not have separate migration data for the British countries. I therefore use the total United Kingdom figure of `r m1[m1$Team=="England" & m1$variable =="Foreign Born Population (%)",3]`% for each of England, Scotland and Wales.] The blue bars repeat the foreign born player percentages from the graph above and the red bars show the percentage of each country’s population that was foreign born. `r WB$Team[1]` has the highest proportion of its population born overseas, at `r WB$value[1]`%, followed by `r WB$Team[2]` with `r WB$value[2]`% and `r WB$Team[3]` with `r WB$value[3]`%.  By comparing each country's respective blue and red bars, we see that Argentina, Australia, France, New Zealand and South Africa have rugby teams which *under*-represent the foreign-born population. Foreign-born players are *over*-represented for every other country.

<br>

<h4>
<center>
Foreign Born Players vs Foreign Born Populations
</center>
</h4>

<center>
```{r, results='asis',echo=FALSE,warning=FALSE}
popcompare_g
```
</center>

<br>

The prevalence of Pacific born players in foreign teams has garnered particular interest with the rugby public and media, although just `r nrow(PI_df)` of the top 10 major countries have had more than two players born in the Pacific Island nations of Samoa, Fiji, and Tonga since 1996.^[While acknowledging the problematic nature of collectively labelling the diverse communities within the South Pacific, for the purposes of this discussion "the Pacific" refers to the countries of Samoa, Fiji and Tonga - the three Pacific Island countries with the largest rugby playing populations.] `r PI_df[PI_df$Team=="Aus",2][[1]]` Wallabies, `r PI_df[PI_df$Team=="Eng",2][[1]]` English players and `r PI_df[PI_df$Team=="NZ",2][[1]]` All Blacks were born in the Pacific. However, these numbers do not tell the full story because each country has vastly different Pacific populations. `r PI_df[PI_df$Team=="NZ",9][[1]]`% of the New Zealand population were born in the Pacific Islands, compared to just `r PI_df[PI_df$Team=="Aus",9][[1]]`% in Australia and less than 0.1% in England.^[These figures are calculated using 2015 [United Nations](http://www.un.org/en/development/desa/population/migration/data/estimates2/data/UN_MigrantStockByOriginAndDestination_2015.xlsx) data for Pacific born populations and [World Bank](http://data.worldbank.org/data-catalog/Population-ranking-table) data (accessed on 22 April 2017) for population counts.] Pacific athletes are wonderful rugby players, and it is unsurprising that they are *over*-represented in all these teams. However, the extent to which they are over represented differs markedly across teams. While New Zealand has `r ceiling(PI_df[PI_df$Team=="NZ",10][[1]])` times more Pacific born players than would be expected given the population, Australia has `r ceiling(PI_df[PI_df$Team=="Aus",10][[1]])` times more, and England a whopping `r ceiling(PI_df[PI_df$Team=="Eng",10][[1]])` times more.

<br>
 
<h4>
<center>
Pacific Born Players vs Pacific Born Populations
</center>
</h4>

<center>
```{r, results='asis',echo=FALSE,warning=FALSE,message=FALSE}
library(cowplot)
plot_grid(PI_g, PIpop_g, PIratio_g, align='h', labels=c('','',''),ncol=3)
```
</center>


The graph below shows the top 10 "suppliers" of players to the rest of the world, which includes all countries for which I have data.^[The next two graphs involving "exports" use data from all 16 countries. This is a little problematic because there are many more missing datapoints. For example, I am missing birthplace data for `r fin_df2$MissPlayers[fin_df2$Team=="Japan"]` Japan players and `r fin_df2$MissPlayers[fin_df2$Team=="USA"]` USA players. The presented numbers therefore represent lower-bound estimates for exports and the trade balance.] `r Expo$Players[1]` `r Expo$Birthcountry[1]`ers have played for another country. The next largest exporter is `r Expo$Birthcountry[2]` with `r Expo$Players[2]` players. Notable among this list are  Samoa, Tonga and Fiji, which together have contributed `r Expo1$Players[Expo1$Birthcountry=="Samoa"]+Expo1$Players[Expo1$Birthcountry=="Fiji"]+Expo1$Players[Expo1$Birthcountry=="Tonga"]` players to the international rugby community. Notable absences include Scotland, Ireland and Wales, which have provided fewer than 7 players each to other countries, despite having had the services of `r FB$ForeignPlayers[FB$Team=="Ireland"]+FB$ForeignPlayers[FB$Team=="Wales"]+FB$ForeignPlayers[FB$Team=="Scotland"]` foreign born players combined.

<br>
<h4>
<center>
Exports of Foreign Players
</center>
</h4>

<center>


```{r, results='asis',echo=FALSE,warning=FALSE}
Expo_g
```
</center>



The next graph summarises the net contribution of each country to the international rugby stock. The "trade balance" is calculated by subtracting the number of "imported" players from the number of "exported" players. A positive trade balance indicates that a country contributes more players to other countries than it "poaches". Leading the way is `r Trade$Country[1]`, which has provided `r Trade[1,4]` more players to other countries than it has poached. On the other side of the ledger is `r Trade$Country[length(Trade$Country)]`, which has imported `r -Trade$"Trade Balance"[Trade$Country=="Scotland"]` more players than it has exported.

<br>

<h4>
<center>
The Trade Balance
</center>
</h4>



<center>
```{r, results='asis',echo=FALSE,warning=FALSE}
Trade_g
```
</center>

<br>

The final table compares native and foreign born contributions on the field. It shows that foreign players tend to score more tries on average than their native born team mates, with the exception of foreign born players in Australia, England and Ireland. However, only Ireland, Italy and South Africa have foreign born players that score more points on average than their native-born counterparts, although these figures are skewed by foreign-born kickers such as Ronan O’Gara and Percy Montgomery.  In terms of number of tests played, there is little difference.

<br>

<h4>
<center>
Player Contributions
</center>
</h4>


```{r, results='asis',echo=FALSE,warning=FALSE}
library(ascii)
print(ascii(cont,include.rownames = F,format='f',digits=c(1),
             align = c("l","c","c","c","c","c","c","c"),
            col.width = c(4,1,1,1,1,1,1,1)),type='pandoc')
```

<br>

<h4 id="data">
<center>
Dataset
</center>
</h4>

You can scroll and search the dataset below, which contains every player from the ten major countries who debuted after 1900. You can download the full dataset [here](https://github.com/hautahi/Rugby-Wanderers/raw/master/data/1.main_data.csv), or you can view each of the individual team datasets at the following links ([Argentina](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Argentina.csv), [England](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/England.csv), [Australia](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Australia.csv), [New Zealand](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/NewZealand.csv), [Scotland](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Scotland.csv), [Ireland](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Ireland.csv), [Wales](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Wales.csv), [South Africa](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/SouthAfrica.csv), [Japan](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Japan.csv), [Samoa](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Samoa.csv), [Fiji](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Fiji.csv), [Tonga](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Tonga.csv),
[Canada](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Canada.csv), [USA](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/USA.csv), [Italy](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/Italy.csv),
[France](https://github.com/hautahi/Rugby-Wanderers/blob/master/data/France.csv)). The Python and R programs that document both the data scraping and adjustments are available at the [Rugby Wanderers Github Repository](https://github.com/hautahi/Rugby-Wanderers).

<br>

```{r, results='asis',echo=FALSE,warning=FALSE}
library(DT)
datatable(han,rownames = FALSE,extensions = c('Scroller'),
          options = list(deferRender = F, scrollY = 200, scroller = TRUE,
                         scrollCollapse=TRUE,scrollX = TRUE,
                         dom = 'Bfrtip',pageLength = 10,
                          columnDefs = list(list(className = 'dt-center', targets = 2:5),list(className = 'dt-left', targets = 0:1),
                                            list(className = 'dt-left', targets = 6))),
          colnames=c('Tests' = 'Matches', 'Pts' = 'Points'))
```

<br>

To my knowledge, the player names and playing statistics are accurate (as at May 15, 2017), but the birthplace information remains a work in progress. The data is especially sparse for Canada, the USA, Argentina and the Pacific Island countries. I welcome any contributions and corrections. Feel free to email me (hautahikingi@gmail.com) with any help you can provide.